---
title: "Download CDEC"
author: "Catarina Pien"
date: "1/29/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Code for downloading data from multiple stations OR multiple sensors on CDEC

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Download Rtools if you don't have latest version at https://cran.r-project.org/bin/windows/Rtools/    
    * Follow instructions above. May need to restart R a few times during the process to get the make folder to show up correctly. 
  
2. Download CDEC Retrieve package

```{r}
devtools::install_github("flowwest/CDECRetrieve", force=TRUE)

# Alt method:
# install roxygen2 (normal install method)
# install.packages("https://cran.r-project.org/src/contrib/Archive/CDECRetrieve/CDECRetrieve_0.1.4.tar.gz", repos=NULL, method = "libcurl")
# Or use winzip to unzip the files, copy folder into the .libPaths() specified (or go to Tools Install to check where lib is)
```

3. Load Packages

```{r setup, include=FALSE}
library(tidyverse)
library(data.table) #rbindlist()
library(lubridate) #today()
library(CDECRetrieve) #cdec_datasets
library(readr)
```


4. Get some metadata on station sensors 

```{r Station sensors, message = FALSE}

# Make a list of the stations you are interested in 
sta_all<-c("BAC", "BET", "BLP", "LIS")

# Filter out list of unique stations (some are in there multiple times)
sta <- sort(unique(sta_all))

### First download dataset info ---------------------------------------------

# Loop to get info on what sensors are available for each station
# Based on station list above
sensordataset <- list() 
for(i in 1:length(sta)){
  dataset <- cdec_datasets(sta[i])
  dataset$station <- sta[i]  # just for an identifier
  sensordataset[[i]] <- dataset
} 

# Bind into dataframe 
cdec_datasets_df <-rbindlist(sensordataset)   #** LIST INTO DATAFRAME **

# Remove unnecessary objects to save space
rm(dataset)
rm(sensordataset)

# Filter to looking at only stations with the sensors you are interested in. 
# E.g. below is only for water temp event(15-min) and hourly data. 
# See cdec_datasets_df for the sensor numbers that are relevant to you
# Sensor 25 = Temp_C
# Sensor 146 = Temp_F
cdec_temp <- cdec_datasets_df %>% 
  filter(sensor_number %in% c(25, 146)) %>% 
  filter(duration !="daily") %>%
  arrange(station) %>%
  mutate(days = end-start) # calculate duration of data

```

5. Download Raw Data from CDEC
These are at Pacific Daylight Time
See https://info.water.ca.gov/misc/sql_dates.html for information about how CDEC deals with Daylight Savings Time

This step may take a while.

To query for multiple stations but one sensor:
```{r Download, message = FALSE}

### Define list of stations ----------------------------------
stations <- c("BLP", "LIS")

# Define start and end dates - these will remain the same throughout
start <- "2018-01-01"
end <- "2018-12-31"

# Define sensor number, time interval 
sensors <- c("25") # Water Temperature Celsius
interval <- "E" # Event = every 15 minutes, H = Hourly, A = Annual

### Download data, bind, write --------------------------------------------
Temp_Event <- lapply(stations, 
                 function(x){
                   cdec_query(station = x,
                              sensor_num = sensors,
                              dur_code = interval,
                              start_date = start,
                              end_date = end)
                 })

Temp_df <- bind_rows(Temp_Event) # bind rows into data frame

```

If you want to query multiple sensors but one station:
```{r}

### Define station----------------------------------
stations <- "LIS"

# Define start and end dates - these will remain the same throughout
start <- "2010-01-01"
end <- "2019-12-31"

# Define sensor number, time interval 
sensors <- c("25", "221") # Water Temperature Celsius and Turbidity
interval <- "E" # Event = every 15 minutes


### Download data, bind, write --------------------------------------------
Data_sensors <- lapply(sensors, 
                 function(x){
                   cdec_query(station = stations,
                              sensor_num = x,
                              dur_code = interval,
                              start_date = start,
                              end_date = end)
                 })

sensors_df <- bind_rows(Data_sensors) # bind rows into data frame
```

Write data 
```{r}
write_csv(sensors_df, "Data/Temp_raw.csv") #write data
saveRDS(sensors_df, "Data/Temp_raw.rds") #write rds
```

