---
title: "InundationPlots"
author: "Catarina Pien"
date: "10/25/2021"
output: html_document
---
# Inundation Plots

    This code plots CPUE data vs. inundation to show the response of CPUE to inundation. Zooplankton data is used as an example.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
Load packages
```{r}
library(tidyverse)
library(readr)
library(lubridate)
```

## Read Inundation dataset and Zoop dataset
```{r}
InundData <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=edi.840.1&entityid=186964642c42e5a8b8e44fc87ff10bbf")
ZoopData <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=edi.494.1&entityid=9190cd46d697e59aca2de678f4ca1c95", col_types = cols(Tide = "f", SpCnd = "d", pH = "d", DO = "d", Turbidity = "d", MicrocystisVisualRank = "f", Comment_PQC = "c"))
glimpse(ZoopData)
```

## Summarize bio data by date
* Filter to species and station and years
* Calculate mean CPUE for each day
* Took out one giant peak to see the plot better

```{r}
ZoopSum <- ZoopData %>%
  filter(StationCode == "STTD", 
         TaxonName %in% c("Cladocera", "Cyclopoida", "Copepoda"),
         WY > 2005)%>%
  group_by(StationCode, Date) %>%
  summarize(meanCPUE = mean(CPUE_ed, na.rm = TRUE)) %>%
  filter(meanCPUE<1e+05) %>%
  ungroup()
```

## Edit inundation data
* Filter to dates of interest
* Assign inundation as occurring up to 10 days after official "YES" for uninundation
* Assign a numerical value to plot in the background for inundation (use max of CPUE, though in this example a high CPUE was removed to better visualize data)
```{r}
inundation <- select(InundData, -c(StageHeight_FW:Overtopping_LIS)) %>%
  filter(Date>"2005-10-01" & Date<"2019-01-01") %>%
  mutate(Inundation2 = ifelse(lead(Inundation, 10) == "TRUE", "TRUE", Inundation),
         Inundation_n = ifelse(Inundation2 == "TRUE", 9e+04, 0))
```

## Make inundation plot
```{r}
# Overall CPUE
ggplot() +
  geom_col(data = inundation, aes(x = Date, y = Inundation_n), fill = "light blue") +
  geom_line(data = ZoopSum, aes(x = Date, y = meanCPUE)) +
  scale_x_date(date_breaks = "6 months") +
  labs(y = "Mean CPUE") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))
```

