---
title: "Mapping_tutorial"
author: "Catarina Pien"
date: "4/2/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Making static and interactive maps {.tabset}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(tidyverse)
library(viridis)
```

Read in data - should have a Latitude and Longitude column in your data
```{r}
stationLatLon <- read.csv("data/CDEC_stations.csv")
stationData <- read.csv("data/StationSensorMetadata.csv") %>%
  stat

stations <- left_join(stationLatLon, stationData, by = c("Station", "StationName")) 
stations <-stations %>% 
  mutate(Duration = 2019-StartYear) %>%
  select(-c(Contact, Email, IncludedInDataset, BackupSensor:DataTreatmentDescription, ChangesSensorDeployment:SensorAccuracy, Comments:EndDateAdditionalData))

# make up some extra variables to visualize
stations$Temp <- runif(nrow(stations), min = 5, max = 50)
```


## Static Maps

### Basic map (Code from Sarah Perry) - https://github.com/seperry83/cat-mentorship/blob/main/map-basics.Rmd

```{r}
library(ggmap)

# Define coordinate bounding box. You could also use numbers if you want. 
buffer = 0.1
coordDict = list( 
    'minLat' = min(stations$Latitude) - buffer,
    'maxLat' = max(stations$Latitude) + buffer,
    'minLon' = min(stations$Longitude) - buffer,
    'maxLon' = max(stations$Longitude) + buffer
)

# Create map object using your bounded coordinates
map_obj <- get_stamenmap(
  bbox = c(left = coordDict[['minLon']], bottom = coordDict[['minLat']], right = coordDict[['maxLon']], top = coordDict[['maxLat']]), # the bounding box
  zoom = 8, # zoom lvl; higher number = more detail (but also more processing power)
  maptype = 'terrain-background'# type of basemap; 'terrain' is my default, but check help(get_stamenmap) for a full list
  )

# Plot the map
map <- ggmap(map_obj, legend = "right")
map

# Add your points
map2 <- map +
  geom_point(data = stations, aes(x = Longitude, y = Latitude, color = Basin.Name), size = 3) +
 geom_text(data = stations, aes(label = Station, x = Longitude, y = Latitude), size = 2, vjust = 0, hjust = 0) +
  labs(x = "Longitude", y = "Latitude")
map2

# Visualize by a value by size or color
map +
  geom_point(data = stations, aes(x = Longitude, y = Latitude, fill = Temp, size = SensorDepth), shape = 21) +
  scale_size(range = c(1,5), limits = c(0,10)) + 
  labs(x = "Longitude", y = "Latitude")

```

Watercolor map example!
Adding gridlines
```{r}
map_obj2 <- get_stamenmap(
  bbox = c(left = coordDict[['minLon']], bottom = coordDict[['minLat']], right = coordDict[['maxLon']], top = coordDict[['maxLat']]), # the bounding box
  zoom = 8, # zoom lvl; higher number = more detail (but also more processing power)
  maptype = 'watercolor'# type of basemap; 'terrain' is my default, but check help(get_stamenmap) for a full list
  )

# Visualize by a value by color
# Add gridlines (probably a better way to do this?)
ggmap(map_obj2, legend = "right")+
  # geom_hline(yintercept =  38.75, color = "darkgrey") +  
  # geom_hline(yintercept =  38.5, color = "darkgrey") +
  #    geom_hline(yintercept =  38.25, color = "darkgrey") +
  # geom_hline(yintercept = 38.0, color = "darkgrey") +
  # geom_hline(yintercept = 37.75, color = "darkgrey") +
  # geom_vline(xintercept = -122.00, color = "darkgrey") + 
  # geom_vline(xintercept = -121.75, color = "darkgrey") + 
  # geom_vline(xintercept = -121.50, color = "darkgrey") +
  # geom_vline(xintercept = -121.25, color = "darkgrey") +
  geom_point(data = stations, aes(x = Longitude, y = Latitude, fill = Duration), shape = 21, size = 3) +
  scale_fill_viridis() + 
  labs(title = "CDEC Station Duration of Data", x = "Longitude", y = "Latitude") +
  theme(panel.grid.major = element_line(color = gray(.5), linetype = "dashed", size = 0.5))
```


### Working with sf package (shapefiles)
https://cengel.github.io/R-spatial/mapping.html

Convert your points to sf object, which gives it a "geometry" column
```{r}
library(sf)# main package for working with shapefiles
# Define the projection of your points, usually WGS 84 (= crs 4326)
stations_sf <- st_as_sf(stations, coords = c("Longitude", "Latitude"), crs = 4326)

# Look at shapefile
head(stations_sf)
plot(stations_sf$geometry)
plot(stations_sf)

```


https://github.com/InteragencyEcologicalProgram/deltamapr

Download a shapefile
```{r}
#install deltamapr
#devtools::install_github("InteragencyEcologicalProgram/deltamapr")
require(deltamapr) # package created by Sam Bashevkin, containing some useful delta shapefiles
require(ggspatial) # north arrows and scale bars
# require(ggsn) alternative for north arrows and scale bars. I didn't like this on eas much.

# Bring this delta waterways shapefile (from deltamapr) into your environment
WW_Delta <- WW_Delta

# Look at file
head(WW_Delta)
plot(WW_Delta)
plot(WW_Delta$geometry)

# Check projection of file
st_crs(WW_Delta)

# Change projection of our stations file to align with Delta waterways file
# Sometimes if you are using multiple files with different projections, they won't match up.
stationsT <- st_transform(stations_sf, crs = 4269) # NAD83 

# Check what classes these files are to make sure they are compatible
class(stationsT)
class(WW_Delta)


```


Make maps!
```{r}
# Plot!
ggplot() +
    geom_sf(data = WW_Delta, fill = "steelblue") +
    geom_sf(data = stationsT, aes(fill = Basin.Name, size = Duration), shape = 21) + 
    scale_fill_viridis(discrete = TRUE) + 
    ggtitle("Duration of Stations in the Delta in Years") +
    theme_bw()

# using EDSM stratum (a type of region)
strMap <- ggplot() +
  geom_sf(data = WW_Delta, fill = "steelblue") + 
  geom_sf(data = R_EDSM_Strata_17P3, aes(fill = Stratum), alpha = 0.4) + 
  geom_sf(data = stationsT, size = 2)+
  scale_fill_brewer(palette = "Dark2") +
  theme_bw()+
  theme(axis.title = element_blank())
strMap

# Add scale bar and north arrow, grid lines
# https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html

  
strMap2 <- strMap + 
  annotation_north_arrow(location = "tr", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
  annotation_scale(location = "bl", bar_cols = c("pink", "white", "pink", "white")) +
  theme(axis.title = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed", size = 0.5))
strMap2 

# Adding text
strMap2 + 
  geom_text(data = stations, aes(x = Longitude, y = Latitude, label = Station), size = 3, check_overlap = FALSE, color = "darkblue", nudge_x = 0.05, nudge_y = 0.05) + 
 annotate(geom = "text", x = -122.4, y = 37.85, label = "San Francisco Bay", fontface = "italic", color = "grey22", size = 3.5 ) 
  
# Add basemap
# https://cengel.github.io/R-spatial/mapping.html

# Redefine your boundaries
coordDict3 = list( 
    'minLat' = 37.25,
    'maxLat' = 39,
    'minLon' = -122.75,
    'maxLon' = -121
)

# Map 
map_obj3 <- get_stamenmap(
  bbox = c(left = coordDict3[['minLon']], bottom = coordDict3[['minLat']], right = coordDict3[['maxLon']], top = coordDict3[['maxLat']]), # the bounding box
  zoom = 9, # zoom lvl; higher number = more detail (but also more processing power)
  maptype = 'terrain-background'# type of basemap; 'terrain' is my default, but check help(get_stamenmap) for a full list
  )

# Plot the map
# Note the inherit.aes = FALSE! Otherwise this does not work.

map_final <- ggmap(map_obj3) + 
  geom_sf(data = WW_Delta, fill = "steelblue", inherit.aes = FALSE) + 
  geom_sf(data = R_EDSM_Strata_17P3, aes(fill = Stratum), alpha = 0.5, inherit.aes = FALSE) + 
  geom_sf(data = stationsT, size = 2, inherit.aes = FALSE)+
  annotation_north_arrow(location = "tr", which_north = "true", 
        pad_x = unit(.1, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
  annotation_scale(location = "bl", bar_cols = c("black", "white", "black", "white")) +
  annotate(geom = "text", x = -122.4, y = 37.85, label = "San Francisco Bay", fontface = "italic", color = "grey22", size = 3.5 ) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Map of CDEC Stations") + 
  theme_bw()+
  theme(axis.title = element_blank()) 
map_final
```


Export image:

https://stackoverflow.com/questions/38907514/saving-a-high-resolution-image-in-r    
jpeg(), tiff(),png() are also functions
```{r, eval = FALSE}
map_final
ggsave("StationMap.jpeg", width = 8, height = 5, device = 'jpeg', dpi = 300)

```

Export shapefiles: 
```{r, eval = FALSE}
st_write(stations_sf, "data/Stations.shp", delete_layer = T, )
st_write(stations_sf, "data/StationsFile.csv", overwrite = T, layer_options = "GEOMETRY=AS_XY")
```


Export map file so you can later load and edit it easily:
```{r, eval = FALSE}
save(map_final, file = "CDEC-station-map.Rdata")
load("CDEC-Station-map.Rdata")
```

Set opacity of basemap: 
https://stackoverflow.com/questions/38126102/set-opacity-of-background-map-with-ggmap


For creating regions on your map
https://github.com/rosehartman/TempRegions/blob/master/app.R











## Interactive Maps 

### Leaflet

Discrete Data Color-coding

Note you can change the label to display what you want. You can also make the labels fancier by writing some html.     
https://rstudio.github.io/leaflet/choropleths.html
```{r}
library(leaflet)
pal <- colorFactor(viridis(7), stations$Basin.Name)
# alternatively, this could be colorNumeric, colorBin, or colorQuantile
# https://rstudio.github.io/leaflet/colors.html

stations %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    color = ~pal(Basin.Name),
    stroke = FALSE,
    fillOpacity = 0.8,
    lng = ~Longitude,
    lat = ~Latitude,
  label = ~paste(Station, "Lat:", Latitude, "Long:", Longitude))%>%
  addLegend(pal = pal,
            values = ~Basin.Name,
            position = "bottomright")

```

Continuous Data change size of dots
```{r}
# You can also manually assign colors to the palettes. 

pal2 <- colorFactor(c("purple3", "blue3", "forestgreen", "orange", "orangered", "hotpink", "grey"), domain = c("CACHE CR", "DELTA", "EAST BAY", "MOKELUMNE R", "SACRAMENTO R", "SAN JOAQUIN R", "STANISLAUS R"))


stations %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    fillColor = ~pal2(Basin.Name),
    radius = ~sqrt(Temp),
    stroke = FALSE, #circle outline
    fillOpacity = 0.6,
    lng = ~Longitude,
    lat = ~Latitude,
   label = ~paste(Station, "Lat:", Latitude, "Long:", Longitude, "Temp:", Temp))%>%
  addLegend(pal = pal2,
            values = ~Basin.Name,
            position = "bottomright")
```


### Mapview

If you hover over, tells you your lat/lon
```{r}
library(mapview)
mapview(stations_sf)
mapview(stations_sf, zcol = "Basin.Name")

```



## Spatial Operations

```{r, eval = FALSE}
st_intersection()
st_buffer()
st_boundary()
```










## Random - I found this interesting! You can use online databases to find shops in Sacramento, or other amenities! 
https://dominicroye.github.io/en/2018/accessing-openstreetmap-data-with-r/

```{r }
library(osmdata)

head(available_features())
head(available_tags("amenity"))
head(available_tags("shop"))

q <- getbb("Sacramento") %>%
      opq() %>%
       add_osm_feature("amenity", "restaurant")


restaurant <- osmdata_sf(q)
restaurant

#our background map
sac_map <- get_map(getbb("Sacramento"), maptype = "toner-background")

#final map
ggmap(sac_map)+
  geom_sf(data = restaurant$osm_points,
          inherit.aes = FALSE,
          colour = "#238443",
          fill = "#004529",
          alpha = .5,
          size = 4,
          shape = 21)+
  labs(x = "", y = "")
```

